### 2014-08-12 Docs/flowchart/flowchart.tex
more equation work

Additionally, we carry species around which can react.  The creation and destruction
of the species is described by their create rate, $\omegadot_k$, and the species
are defined by their mass fractions, $X_k \equiv \rho_k / \rho$, giving
\begin{equation}
\frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub X_k) = \rho \omegadot_k
\end{equation}
and
\begin{equation}
\sum_k X_k = 1
\end{equation}

@@

As first discussed in paper III and then refined in the multilevel paper, we capture
the changes that can occur due to significant convective
overturning by imposing the constraint that $\overline{\rho'}=0$
for all time.  This gives
\begin{equation}
\frac{\partial\overline{\rho'}}{\partial t} = -\nabla\cdot(\etarho\eb_r).
\end{equation}
where
\begin{equation}
\etarho = \overline{\left(\rho'\Ub\cdot\eb_r\right)}
\end{equation}
In practice, we correct the drift by simply setting $\rho_0 =
\overline{\rho}$ after the advective update of $\rho$.  However we still need to
explicitly compute $\etarho$ since it appears in other equations.

@@

The equation of state is cast into an elliptic constraint on the
velocity field by differentiating $p_0(\rho, s, X_k)$ along particle
paths, giving:
\begin{equation}
\nabla \cdot (\beta_0 \Ub) = 
   \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
\label{eq:U divergence}
\end{equation}
where $\beta_0$ is a density-like variable that carries background
 stratification, defined as
\begin{equation}
\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
\end{equation}
and 
\begin{equation}
S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext + \frac{\sigma}{\rho} \nabla \cdot \kth \nabla T
\end{equation}
where $p_{X_k} \equiv \left. \partial p / \partial X_k
\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h / 
\partial X_k \right |_{p,T,X_{j,j\ne k}}, 
p_\rho \equiv \left. \partial p/\partial \rho \right |_{T, X_k}$, and
$\sigma \equiv p_T/(\rho c_p p_\rho)$, with
$p_T \equiv \left. \partial p / \partial T \right|_{\rho, X_k}$ and
$c_p \equiv \left.  \partial h / \partial T
\right|_{p,X_k}$ is the specific heat at constant pressure.   The last
term is only present if we are using thermal diffusion (\runparam{use\_thermal\_diffusion} {\tt = T}).  In this term, $\kth$ is the thermal conductivity.

@@

We decompose the full velocity field into a base state velocity,
$w_0$, that governs the base state dynamics, and a local velocity,
$\Ubt$, that governs the local dynamics, i.e., 
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$---the motivation for this splitting was given in paper~II.
The velocity evolution equations are then 
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r},\label{eq:w0
  evolution}\\ 
%
\frac{\partial\Ubt}{\partial t} &=& -(\Ubt + w_0\er)\cdot\nabla\Ubt
  - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = 
    \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}
           \frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting
(\ref{eq:w0 divergence}) into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

@@

The base state evolution equations for density and enthalpy can be
found by averaging Eq~\ref{eq:flow:enthalpy}
over a layer of constant radius, resulting in
\begin{eqnarray}
\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot\left[(\rho h)_0w_0\eb_r\right] +
  \psi + \overline{\rho \Hnuc} + \overline{\rho \Hext}. \label{eq:flow:enthalpy_base}
\end{eqnarray}
Subtracting it from the full enthalpy equation gives:
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - (\rho h)'\nabla\cdot\Ub - 
  \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
   + ( \rho\Hnuc - \overline{\rho \Hnuc}) + (\rho\Hext - \overline{\rho \Hext}) 
\end{equation}

@@

\begin{description}

@@

\item Compute $S^{\nph,\star\star}$.  We compute an estimate for the
  time-centered expansion term in the velocity divergence constraint,
  as given in paper~III equation (19),
\begin{equation}
  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip .
\label{eq:defineS} 
\end{equation}

  For the first time step ($n=0$), we set
\begin{equation}
S^{n+\myhalf,\star\star} = \frac{S^0 + S^1}{2},
\end{equation}
  where $S^1$ is found during initialization.  For other time steps
  $(n \ne 0)$, following \cite{SNe}, we extrapolate
  to the half-time using $S$ at the previous and current
  time levels
\begin{equation}
S^{\nph,\star\star} = S^n + \frac{\dt^n}{2} \frac{S^n - S^{n-1}}{\dt^{n-1}}.
\end{equation}
  Next, compute
\begin{equation}
\overline{S^{\nph,\star\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star\star}\right).
\end{equation}

@@

\item Compute the provisional base state velocity forcing, using equation (38)
  from paper~III,
\begin{equation}
-\frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r}, \label{eq:pizero}
\end{equation}
  with the following discretization:
\begin{equation}
\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^{n,\star} = -\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} - w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star},
\end{equation} 
  where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are defined as
\begin{eqnarray}
w_0^{n,\star} &=& \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}, \\
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} &=& \frac{1}{\dt^n+\dt^{n-1}}\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh} + \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ].\nonumber \\
\end{eqnarray}
  If $n=0$, we use $\dt^{-1} = \dt^0$.
\end{enumerate}

@@

The local velocity field is described by paper III equation (37),
\begin{equation}
\frac{\partial\Ubt}{\partial t} = 
- \left(\Ubt+w_0\right) \cdot \nabla \Ubt
- \left(\Ubt \cdot \eb_r\right) \frac{\partial w_0}{\partial r} \eb_r
- \frac{1}{\rho} \nabla\pi
+ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \eb_r
- \frac{(\rho-\rhozero)}{\rho} \; g \; \eb_r  \label{eq:utildeupd}  \enskip .
\end{equation}
From this, we compute time-centered edge 
velocities, $\uadvonedag$, using 
$\Ub = \Ubt^n + w_0^{\nph,\star}$.  The $\dagger$ superscript refers to the 
fact that the predicted velocity field does not satisfy the divergence 
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection, as described in detail in Appendix B of Paper III.  
We note that $\uadvone$ satisfies the discrete version of
$\overline{(\uadvone\cdot\eb_r)}=0$ as well as
\begin{eqnarray}
\nabla \cdot \left(\beta_0^n \uadvone\right) &=& \beta_0^n \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right),\\
 \beta_0^n &=& \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma_1^n}\right),
\end{eqnarray}
where $\beta_0$ is computed as described in Appendix C of Paper III.

@@

\item Update $(\rho X_k)$ using a discretized version of the species 
continuity equation,
\begin{equation}
\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\Ub \rho X_k) = 0 \enskip .
\label{eq:species}
\end{equation}
where we omit the reaction terms, which were already 
accounted for in {\bf React State}.  The update consists of two steps:

@@

  There are a variety of choices of quantities to predict to the
  edges (controlled by \runparam{species\_pred\_type}---see Chapter~\ref{ch:pert}).
  By default, we use the equations 
\begin{eqnarray}
\frac{\partial\rho'}{\partial t} &=& -\Ub\cdot\nabla\rho' - 
     \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right),
     \label{eq:Perturbational Density}  \\
\frac{\partial X_k}{\partial t} &=& -\Ub\cdot\nabla X_k + 
     \omegadot_k. \label{eq:Primitive Species}
\end{eqnarray}
  to
  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
  and $X_k^{\nph,\star,\pred}$.
  We convert the perturbational density to full state density using
\begin{equation}
\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
\end{equation}
  where the base state density terms are mapped to Cartesian edges.
  Then,\\
  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.

@@

\item Define a radial edge-centered $\etarho^{\nph,\star}$.

@@

\item Update the enthalpy using a discretized version of the enthalpy
evolution equation, again omitting the reaction and heating terms
since we already accounted for
them in {\bf React State}.  This equation takes the form:
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r}.
\end{equation}
For spherical geometry, we solve the
analytically equivalent form,
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt,
\end{equation}
which experience has shown to minimize the drift from thermodynamic
equilibrium.  The update consists of two steps:

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.  There are a variety
   of quantities that we can predict to the interfaces here (controlled
   by \runparam{enthalpy\_pred\_type}---see Chapter~\ref{ch:pert}).  For
   the default case, we use the perturbational
  enthalpy equation, neglecting reactions,
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - 
   (\rho h)'\nabla\cdot\Ub - \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
    \label{eq:Perturbational Enthalpy}.
\end{equation}
  to predict
  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
  enthalpy to a full state enthalpy using
\begin{equation}
(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
\end{equation}
  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
  In spherical geometry, our experience has shown that a slightly different
  approach leads to reduced discretization errors.  We first map 
  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
  and then multiply these terms to get $(\rho h)_0$.

@@

Then, for each Cartesian cell where $\rho^{(2),\star} < \rho_\mathrm{cutoff}$,
we recompute enthalpy using
\begin{equation}
(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
\end{equation}

\item Update the temperature using the equation of state:
$T^{(2),\star} = 
  T(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star})$ (planar geometry) or
$T^{(2),\star} = 
  T(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star})$ (spherical geometry).
\end{enumerate}

@@

\item Compute $S^{\nph,\star}$.  First, compute $S^{n+1,\star}$ with
\begin{equation}
S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
\end{equation}
  where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} /
  \rho^{(2),\star}$ and the thermodynamic quantities are defined using
  $\rho^{n+1,\star}, X_k^{n+1,\star},$ and $T^{n+1,\star}$ as inputs to
  the equation of state.  Then, define
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} (S^{\nph,\star}),
\qquad
 S^{\nph.\star} = \frac{S^n + S^{n+1,\star}}{2},
\end{equation}
\item Compute $w_0^{\nph}$.  First, define
\begin{equation}
\overline{\Gamma_1^{\nph,\star}} = \frac{\overline{\Gamma_1^n} + \overline{\Gamma_1^{n+1,\star}}}{2}, 
\quad
\rho_0^{\nph,\star} = \frac{\rho_0^{n} + \rho_0^{n+1,\star}}{2},
\quad
p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2},
\end{equation}
  with
\begin{equation}
 \overline{\Gamma_1^{n+1,\star}} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right].
\end{equation}
\begin{description}
\item For planar geometry, call\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},p_0^{\nph,\star},\psi^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
\item For spherical geometry, call\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},\rho_0^{\nph,\star},p_0^{\nph,\star},\etarho^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
\end{description}

>>>>

Additionally, we carry species around which can react.  The creation and destruction
of the species is described by their create rate, $\omegadot_k$, and the species
are defined by their mass fractions, $X_k \equiv \rho_k / \rho$, giving
\begin{equation}
\frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub X_k) = \rho \omegadot_k
\label{eq:flow:rhoX}
\end{equation}
and
\begin{equation}
\sum_k X_k = 1
\end{equation}

@@

As first discussed in paper III and then refined in the multilevel paper, we capture
the changes that can occur due to significant convective
overturning by imposing the constraint that $\overline{\rho'}=0$
for all time.  This gives
\begin{equation}
\frac{\partial\overline{\rho'}}{\partial t} = -\nabla\cdot(\etarho\eb_r).
\end{equation}
where
\begin{equation}
\etarho = \overline{\left(\rho'\Ub\cdot\eb_r\right)}
\label{eq:flow:etarho}
\end{equation}
In practice, we correct the drift by simply setting $\rho_0 =
\overline{\rho}$ after the advective update of $\rho$.  However we still need to
explicitly compute $\etarho$ since it appears in other equations.

@@

The equation of state is cast into an elliptic constraint on the
velocity field by differentiating $p_0(\rho, s, X_k)$ along particle
paths, giving:
\begin{equation}
\nabla \cdot (\beta_0 \Ub) = 
   \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
\label{eq:U divergence}
\end{equation}
where $\beta_0$ is a density-like variable that carries background
 stratification, defined as
\begin{equation}
\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
\end{equation}
and 
\begin{equation}
S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext + \frac{\sigma}{\rho} \nabla \cdot \kth \nabla T
\label{eq:flow:S}
\end{equation}
where $p_{X_k} \equiv \left. \partial p / \partial X_k
\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h / 
\partial X_k \right |_{p,T,X_{j,j\ne k}}, 
p_\rho \equiv \left. \partial p/\partial \rho \right |_{T, X_k}$, and
$\sigma \equiv p_T/(\rho c_p p_\rho)$, with
$p_T \equiv \left. \partial p / \partial T \right|_{\rho, X_k}$ and
$c_p \equiv \left.  \partial h / \partial T
\right|_{p,X_k}$ is the specific heat at constant pressure.   The last
term is only present if we are using thermal diffusion (\runparam{use\_thermal\_diffusion} {\tt = T}).  In this term, $\kth$ is the thermal conductivity.

@@

We decompose the full velocity field into a base state velocity,
$w_0$, that governs the base state dynamics, and a local velocity,
$\Ubt$, that governs the local dynamics, i.e., 
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$---the motivation for this splitting was given in paper~II.
The velocity evolution equations are then 
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r},\label{eq:w0
  evolution}\\ 
%
\frac{\partial\Ubt}{\partial t} &=& -(\Ubt + w_0\er)\cdot\nabla\Ubt
  - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:flow:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = 
    \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}
           \frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting
(\ref{eq:w0 divergence}) into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

@@

The base state evolution equations for density and enthalpy can be
found by averaging Eq~\ref{eq:flow:enthalpy}
over a layer of constant radius, resulting in
\begin{eqnarray}
\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot\left[(\rho h)_0w_0\eb_r\right] +
  \psi + \overline{\rho \Hnuc} + \overline{\rho \Hext}. \label{eq:flow:enthalpy_base}
\end{eqnarray}
Subtracting it from the full enthalpy equation gives:
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - (\rho h)'\nabla\cdot\Ub - 
  \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
   + ( \rho\Hnuc - \overline{\rho \Hnuc}) + (\rho\Hext - \overline{\rho \Hext}) 
\label{eq:flow:rhohprime}
\end{equation}

@@

\begin{description}\itemsep 1em

@@

\item Compute $S^{\nph,\star\star}$.  We compute an estimate for the
  time-centered expansion term in the velocity divergence constraint,
  as given in Eq.~\ref{eq:flow:S}.  For the first time step ($n=0$),
  we set
\begin{equation}
S^{n+\myhalf,\star\star} = \frac{S^0 + S^1}{2},
\end{equation}
  where $S^1$ is found during initialization.  For other time steps
  $(n \ne 0)$, following \cite{SNe}, we extrapolate
  to the half-time using $S$ at the previous and current
  time levels
\begin{equation}
S^{\nph,\star\star} = S^n + \frac{\dt^n}{2} \frac{S^n - S^{n-1}}{\dt^{n-1}}.
\end{equation}
  Next, compute
\begin{equation}
\overline{S^{\nph,\star\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star\star}\right).
\end{equation}

@@

\item Compute the provisional base state velocity forcing, using equation (38)
  from paper~III,
\begin{equation}
-\frac{\beta_0}{\rho_0} \frac{\partial (\pi_0/\beta_0)}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r}, \label{eq:pizero}
\end{equation}
  with the following discretization:
\begin{equation}
\left ( \frac{\beta_0}{\rho_0} \frac{\partial (\pi_0/\beta_0)}{\partial r} \right )^{n,\star} = -\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} - w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star},
\end{equation} 
  where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are defined as
\begin{eqnarray}
w_0^{n,\star} &=& \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}, \\
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} &=& \frac{1}{\dt^n+\dt^{n-1}}\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh} + \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ].\nonumber \\
\end{eqnarray}
  If $n=0$, we use $\dt^{-1} = \dt^0$.
\end{enumerate}

@@

The local velocity field is described by Eq.~\ref{eq:flow:utildeupd}.
From this, we compute time-centered edge 
velocities, $\uadvonedag$, using 
$\Ub = \Ubt^n + w_0^{\nph,\star}$.  The $\dagger$ superscript refers to the 
fact that the predicted velocity field does not satisfy the divergence 
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection, as described in detail in Appendix B of Paper III.  
We note that $\uadvone$ satisfies the discrete version of
$\overline{(\uadvone\cdot\eb_r)}=0$ as well as
\begin{eqnarray}
\nabla \cdot \left(\beta_0^n \uadvone\right) &=& \beta_0^n \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right),\\
 \beta_0^n &=& \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma_1^n}\right),
\end{eqnarray}
where $\beta_0$ is computed as described in Appendix C of Paper III (although
note that an alternate procedure that uses linear reconstruction of $g$ in
constructing $\beta_0$ is enabled with \runparam{use\_linear\_grav\_in\_beta}).

@@

\item Update $(\rho X_k)$ using a discretized version of the species 
continuity equation, Eq.~\ref{eq:flow:rhoX}, where we omit the
reaction terms, which were already accounted for in {\bf React State}:
\begin{equation}
\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\Ub \rho X_k) = 0 \enskip .
\label{eq:species}
\end{equation}
The update consists of two steps:

@@

  There are a variety of choices of quantities to predict to the
  edges (controlled by \runparam{species\_pred\_type}---see Chapter~\ref{ch:pert}).
  By default, we use the equations 
\begin{eqnarray}
\frac{\partial\rho'}{\partial t} &=& -\Ub\cdot\nabla\rho' - 
     \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right),
     \label{eq:Perturbational Density}  \\
\frac{\partial X_k}{\partial t} &=& -\Ub\cdot\nabla X_k + 
     \omegadot_k. \label{eq:Primitive Species}
\end{eqnarray}
  to
  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
  and $X_k^{\nph,\star,\pred}$.
  We convert the perturbational density to full state density using
\begin{equation}
\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
\end{equation}
  where the base state density terms are mapped to Cartesian edges.
  Then,\\
  $(\rho X_k)^{\nph,\star,\pred} = \rho^{\nph,\star,\pred} \, X_k^{\nph,\star,\pred}$.

@@

\item Define a radial edge-centered $\etarho^{\nph,\star}$ (Eq.~\ref{eq:flow:etarho}).

@@

\item Update the enthalpy using a discretized version of the enthalpy
evolution equation (Eq.~\ref{eq:flow:enthalpy}), again omitting the reaction and heating terms
since we already accounted for
them in {\bf React State}.  This equation takes the form:
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r}.
\end{equation}
For spherical geometry, we solve the
analytically equivalent form,
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt,
\end{equation}
which experience has shown to minimize the drift from thermodynamic
equilibrium.  The update consists of two steps:

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.  There are a
   variety of quantities that we can predict to the interfaces here
   (controlled by \runparam{enthalpy\_pred\_type}---see
   Chapter~\ref{ch:pert}).  For the default case, we use the
   perturbational enthalpy equation, Eq.~\ref{eq:flow:rhohprime}, neglecting reactions,
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - 
   (\rho h)'\nabla\cdot\Ub - \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
    \label{eq:Perturbational Enthalpy}.
\end{equation}
  to predict
  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
  enthalpy to a full state enthalpy using
\begin{equation}
(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
\end{equation}
  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
  In spherical geometry, our experience has shown that a slightly different
  approach leads to reduced discretization errors.  We first map 
  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
  and then multiply these terms to get $(\rho h)_0$.

@@

Then, for each Cartesian cell where $\rho^{(2),\star} < \rho_\mathrm{cutoff}$,
we recompute enthalpy using
\begin{equation}
(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
\end{equation}
This behavior is controlled by \runparam{do\_eos\_h\_above\_cutoff}.

\item Update the temperature using the equation of state:
$T^{(2),\star} = 
  T(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star})$ (planar geometry) or
$T^{(2),\star} = 
  T(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star})$ (spherical geometry).
\end{enumerate}
As before, this behavior is controlled by \runparam{use\_tfromp}.

@@

\item Compute $S^{\nph,\star}$.  First, compute $S^{n+1,\star}$ with
\begin{equation}
S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + 
   \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + 
   \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
\end{equation}
  where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} /
  \rho^{(2),\star}$ and the thermodynamic quantities are defined using
  $\rho^{n+1,\star}, X_k^{n+1,\star},$ and $T^{n+1,\star}$ as inputs to
  the equation of state.  If we are using diffusion then we would include the diffusion
  term in $S$.  Then, define
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} (S^{\nph,\star}),
\qquad
 S^{\nph.\star} = \frac{S^n + S^{n+1,\star}}{2},
\end{equation}
\item Compute $w_0^{\nph}$.  First, define
\begin{equation}
\overline{\Gamma_1^{\nph,\star}} = \frac{\overline{\Gamma_1^n} + \overline{\Gamma_1^{n+1,\star}}}{2}, 
\quad
\rho_0^{\nph,\star} = \frac{\rho_0^{n} + \rho_0^{n+1,\star}}{2},
\quad
p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2},
\end{equation}
  with
\begin{equation}
 \overline{\Gamma_1^{n+1,\star}} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right].
\end{equation}
\begin{description}
\item For planar geometry, call\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},p_0^{\nph,\star},\psi^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
\item For spherical geometry, call\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},\rho_0^{\nph,\star},p_0^{\nph,\star},\etarho^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
\end{description}
